\documentclass[12pt, english, reqno]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Stylistic points.
\usepackage{geometry}
\geometry{
  a4paper,
  right=2.5cm,
  left=2.5cm,
  top=2.5cm,
  bottom=2.5cm,
  %showframe,
}

\usepackage{mathpazo} % Math Palatino font.

% Hyperlinks.
\usepackage{xcolor}
\definecolor{Myblue}{rgb}{0,0,0.6}
\usepackage[a4paper,colorlinks,citecolor=Myblue,linkcolor=Myblue,urlcolor=Myblue,pdfpagemode=None]{hyperref}

% Math-related things.
\usepackage{mathtools}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{mathrsfs}
\usepackage[margin=1cm]{caption}
\usepackage{enumitem}
\usepackage{newunicodechar}

% Tikz setup.
\usepackage{tikz}
\usetikzlibrary{cd, arrows, matrix}

% Amsmath stuff.
% \theoremstyle{definition}

\newtheoremstyle{myteo}{\topsep}{\topsep}
	{}
	{}
	{\bfseries}
	{.}
	{2pt}
	{\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}
\theoremstyle{myteo}

\newtheorem{definition}{Definition}[section]
\newtheorem{example}[definition]{Example}
\newtheorem{notation}[definition]{Notation}
\newtheorem{remark}[definition]{Remark}
\newtheorem{fact}[definition]{Fact}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{lemma}[definition]{Lemma}

% Radom stuff.
\usepackage{csquotes}

% Custom commands.
\newcommand{\cat}[1]{{\mathscr #1}}
\newcommand{\catname}[1]{\text{\normalfont\textbf{#1}}}
\newcommand{\vvector}[2]{\begin{pmatrix}#1\\ #2\end{pmatrix}}
\newcommand{\hvector}[2]{\begin{pmatrix}#1 & #2\end{pmatrix}}
\newcommand{\vvectorfour}[4]{\begin{pmatrix}#1\\ #2\\ #3\\ #4\end{pmatrix}}
\newcommand{\vvectormini}[2]{\left(\begin{smallmatrix}#1\\ #2\end{smallmatrix}\right)}
\newcommand{\hvectormini}[2]{\left(\begin{smallmatrix}#1 & #2\end{smallmatrix}\right)}
\newcommand{\fun}[3]{#1\colon #2\to #3}
\newcommand{\epi}[3]{#1\colon #2\twoheadrightarrow #3}
\newcommand{\nat}[3]{#1\colon #2\Rightarrow #3}
\newcommand{\add}[2]{\text{\normalfont Add}[#1, #2]}
\newcommand{\functdef}[7]{\begin{align*}
                            #1\colon #2 &\longrightarrow #3\\
                            #4 &\longmapsto #5\\
                            #6 &\longmapsto #7
                          \end{align*}}
\newcommand{\fundef}[5]{\begin{align*}
                             #1\colon\begin{tikzcd}[sep=large]#2\end{tikzcd} &\longrightarrow \begin{tikzcd}[sep=large]#3\end{tikzcd}\\
                             \begin{tikzcd}[sep=large]#4\end{tikzcd} &\longmapsto \begin{tikzcd}[sep=large]#5\end{tikzcd}
                           \end{align*}}
\newcommand{\homset}[3]{#1(#2,#3)}
\newcommand{\precomp}[1]{{#1}^*}
\newcommand{\postcomp}[1]{{#1}_*}
\newcommand{\nathom}[2]{\text{\normalfont Nat}(#1,#2)}
\newcommand{\eqtext}[1]{\stackrel{\mathclap{\scriptsize\mbox{#1}}}{=}}
\newcommand{\op}[1]{#1^{\text{op}}}
\newcommand{\hh}[1]{{}_={#1}}
\newcommand{\vh}[1]{{#1}^{\parallel}}
\newcommand{\rec}[1]{{}^{\Box}{#1}}
\newcommand{\don}[1]{{#1}_{\Box}}

% Custom operators.
\DeclareMathOperator\coker{coker}
\DeclareMathOperator\im{im}
\DeclareMathOperator*\colim{colim}

% Yoneda.
\newcommand{\yo}{\text{\usefont{U}{min}{m}{n}\symbol{'210}}}
\DeclareFontFamily{U}{min}{}
\DeclareFontShape{U}{min}{m}{n}{<-> udmj30}{}

% Figures stuff.
\renewcommand{\figurename}{Diagram}
\renewcommand{\thefigure}{\arabic{section}.\arabic{figure}}

\begin{document}
% \rhead{\thepage}
% \lhead{\nouppercase{\leftmark}} 

\title{Notes on Abelian Categories and Mitchell's Embedding Theorem}
\author{Gabriele Rastello}
\maketitle
\tableofcontents

\newpage
\section*{Preface}
% These notes are a digested read of (a part of) the first chapter of Borceux's Handbook of Categorical Algebra 2.
% The intention is that of introducing Abelian Categories and then prove the celebrated Mitchell's Embedding Theorem (see \ref{teo:mitchell}).

% The text is tailored toward beginners in Category Theory

\medskip\noindent\emph{Why these notes exist.}
Around September 2019, given my interest in Category Theory and my upcoming graduation, I asked a professor that works in this branch to be the advisor of my graduation thesis.
After a brief meeting he assigned me the first chapter of Borceux's Handbook of Categorical Algebra (see \cite{handbook2}), with Mitchell's Embdedding Theorem as a specific goal.
With this in mind I started reading the book and writing these notes, adding here and there some details to the proofs, examples and so on.
Then the Covid19 pandemic hit and this document became the major way of communicating the progress I was making to my advisor.
Eventually everything was approved and these notes were recombined into my undergraduate thesis.
But in the process of doing so I realized that maybe someone, somewhere, could benefit from reading them and that publishing them wasn't hard at all (we have the internet) so this document was born.

\medskip\noindent\emph{Who could benefit from reading these notes?}
Anyone interested in learning the basics of Abelian Categories and seeing a proof of Mitchell's Embedding Theorem.
As I've already said these notes contain more detailed proofs that \cite{handbook2} and so I feel that can be most useful to beginners in the study of Category Theory.
Moreover they are more self-contained that \cite{handbook2} since most of the needed theorems that Borceux imports from his other book are re-stated and proven as well.
This removes the annoyance of having to continually switch between books.
However a minimal subset of results is not re-stated and re-proven; this mostly have to do with my advisor not requesting an explicit proof of them.
I might decide to include them in a later version of the notes, but this is very unlikely since it will basically require to re-write much of \cite{handbook1}'s chapter 2; that is very accessible to beginners anyway.

\medskip\noindent\emph{What do you mean with the term ``beginners''?}
Basically anyone that has the same degree of knowledge of category theory that I had back in September 2019.
Namely people that are comfortable with the notions of: categories, functors, natural transformations and (co)limits.
In this last requirement I include the definition of (co)limits and all the standard ones (initial/terminal objects, products/coproducs, pullbacks/pushouts, equalizers/coequalizers).
Moreover one should know what groups, abelian groups, rings and modules are, but nothing more than the mere definitions are needed.

\medskip\noindent\emph{A word on section 6.}
The sixth and last section was initially intended to be a brief example of application of Mitchell's Embdedding Theorem.
However it grew into a quick introduction to the homology of (double) complexes followed by a proof of some classical lemmas.
The reader can safely ignore this section. 

\input{sections/preadditive_and_additive_categories.tex}
\input{sections/abelian_categories.tex}
\input{sections/additive_functors.tex}
\input{sections/exact_sequences_and_exact_functors.tex}
\input{sections/the_embedding_theorem.tex}
\input{sections/salamanders.tex}

\newpage
\input{bibliography.tex}
\end{document}
